Analytics Without Parameter Tuning Considered Harmful?

### 2015-03-14 dm101.tex
Updates from ShareLaTeX
safey save

Since programming inherently
introduces defects into  programs, is important to thoroughly {\em test} software before it is {\em used}.
Such testing  can be very expensive.
Software assessment budgets are finite
while assessment effectiveness increases 
exponentially with assessment effort.
Lowry et al. warn that  
the state space explosion problem imposes
strict limits on how much a system can be explored
via automatic formal methods~\cite{lowrey98}.
For standard black-box testing methods,
a {\em linear} increase
in the confidence $C$ that we have found all defects
can take {\em exponentially} more effort.
For example, for one-in-a-thousand detects,
moving $C$ from  
90\% to 98\% nearly doubles the number of  tests (2301 to   3910 black box
probes, respectively)\footnote{A randomly selected 
input to a program will find a fault with probability $p$.
After $N$ random black-box tests, the chances of the inputs 
not revealing any fault 
is $(1-p)^N$. Hence, the chances $C$ of seeing the fault is $1-(1-p)^N$
which can be rearranged to 
 $N(C,p)=\frac{log(1 -
C)}{log(1-p)}$. For example, $N(0.90,10^{-3})=2301$.}.

@@

One such lightweight sampling policy is defect predictors learned from static code attributes.
Given software described using (for example) the attributes of \fig{ck} it is possible
to use standard data mining technology to learn where the probability of software defects is highest.
Such defect  defect predictors learned from
static code attributes are   {\em easy to
use}, {\em widely-used}, and {\em useful} to use.

>>>>

Since programming inherently
introduces defects into  programs, is important to thoroughly {\em test} software before it is {\em used}.
Such testing  can be very expensive.
Software assessment budgets are finite
while assessment effectiveness increases 
exponentially with assessment effort.
Lowry et al. warn that  
the state space explosion problem imposes
strict limits on how much a system can be explored
via automatic formal methods~\cite{lowrey98}.
For standard black-box testing methods,
a {\em linear} increase
in the confidence $C$ that we have found all defects
can take {\em exponentially} more effort.
For example, for one-in-a-thousand detects,
moving $C$ from  
90\% to 98\% nearly doubles the number of  tests (2301 to   3910 black box
probes, respectively)\footnote{A randomly selected 
input to a program will find a fault with probability $p$.
After $N$ random black-box tests, the chances of the inputs 
not revealing any fault 
is $(1-p)^N$. Hence, the chances $C$ of seeing the fault is $1-(1-p)^N$
which can be rearranged to 
 $N(C,p)=log(1 -
C)/log(1-p)$. For example, $N(0.90,10^{-3})=2301$.}.

@@

One such lightweight sampling policy is defect predictors learned from static code attributes.
Given software described using (for example) the attributes of \fig{ck}, it is possible
to use standard data mining technology to learn where the probability of software defects is highest.
Such defect  defect predictors learned from
static code attributes are   {\em easy to
use}, {\em widely-used}, and {\em useful} to use.

### 2015-03-14 learners.tex
safey save

	\end{tabular}
    \caption {List of parameters to be tuned in Where-based learner and CART in scikit-learn.}
\label{fig:parameters}
\end{figure*}

@@

Optionally, if the {\em where prune} option is set, 
WHERE  continues to applies infogain criteria  recursively to build a tree that selects for the
different clusters. If WHERE's {\em tree prune} parameter is enabled, then WHERE also prunes  superfluous sub-trees. For example, if a sub-tree and its parent have the same 
majority cluster
(one that occurs most frequently), then we prune the sub-tree.
This tree pruning  sometimes
prunes aways all  cluster selectors branches. To tame this effect, the {\em wriggle} parameter
blocks tree pruning for at least the the {\em wriggle} number of initial branches.

@@

To find our ``scout'',  we sorted  the above algorithms from simplest to most complex.
The three simplest optimizers are SA, $\mathcal{E}$-domination, and 
differential evolution (each can be coded in less than a page of some high-level scripting language). Our reading of the current literature is that there are more current advocated for
differential evolution than
  SA or $\mathcal{E}$-domination. when the MOEA/D community select a secondary optimizer, then often use  differential evolution~\cite{zhang07,5583335}.
  Also, Vesterstrom and Thomsen~\cite{Vesterstrom04} found that  differential evolution was competitive with 
   particle swarm optimization and a genetic algorithm.

>>>>

	\end{tabular}
    \caption {List of parameters to be tuned.}
\label{fig:parameters}
\end{figure*}

@@

Optionally, if the {\em where prune} option is set, 
WHERE  continues to applies infogain criteria  recursively to build a tree that selects for the
different clusters. If WHERE's {\em tree prune} parameter is enabled, then WHERE also prunes  superfluous sub-trees. For example, if a sub-tree and its parent have the same 
majority cluster
(one that occurs most frequently), then we prune the sub-tree.
This tree pruning  sometimes
prunes aways all  cluster selectors branches. To tame this effect, the {\em wriggle} parameter
blocks tree pruning for at least the first {\em wriggle} number of initial branches.

@@

To find our ``scout'',  we used engineering judgement to sort  the above algorithms from simplest to most complex.
The three simplest optimizers are SA, $\mathcal{E}$-domination, and 
differential evolution (each can be coded in less than a page of some high-level scripting language). Our reading of the current literature is that there are more  advocates for
differential evolution than
  SA or $\mathcal{E}$-domination:
  \bi
  \item
  When the MOEA/D community requires a secondary optimizer, then often use  differential evolution~\cite{zhang07,5583335}.
  \item
 Vesterstrom and Thomsen~\cite{Vesterstrom04} report that DE is competitive with 
   particle swarm optimization and a genetic algorithm. 
   \ei
DEs have been applied before for   parameter tuning (e.g. see~\cite{omran2005differential, chiha2012tuning}) but this is the first time they have been applied to
optimizing defect prediction from static code attributes.

### 2015-03-14 tuningpredictor.tex
safey save

\documentclass{sig-alternative}
\usepackage{multirow}
\usepackage{color}
\usepackage{graphics}
\usepackage{cite}
\usepackage{rotating}
\usepackage{graphics}
\usepackage{colortbl} 
\usepackage{picture}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\renewcommand{\footnotesize}{\scriptsize}
\definecolor{lightgray}{gray}{0.8}
\definecolor{darkgray}{gray}{0.6}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
%%% graph
\newcommand{\crule}[3][darkgray]{\textcolor{#1}{\rule{#2}{#3}}}
\newcommand{\rone}{\crule{1mm}{1.95mm}}
\newcommand{\rtwo}{\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}}
\newcommand{\rthree}{\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}}
\newcommand{\rfour}{\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}} 
\newcommand{\rfive}{\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}}
\newcommand{\quart}[3]{\begin{picture}(100,6)%1
{\color{black}\put(#3,3){\circle*{4}}\put(#1,3){\line(1,0){#2}}}\end{picture}}
\definecolor{Gray}{gray}{0.95}
\definecolor{LightGray}{gray}{0.975}

@@

For the purposes of learning
software defect predictors this optimization strategy can quickly
find  good tunings that  dramatically change   the performance of a learner.
For example,
in this paper we show   tunings that  alter deection  precision  
 from 2\% to 98\%.

These results prompt for a change to standard methods in software analytics.
At least for defect prediction, 
it is no longer enough to just run a data miner and present the result
{\em without} first conducting an tuning optimizations study.
The implications for other kinds of software analytics is now an open and pressing questions.

@@

In the $21^{st}$ century, it is now impossible
to manually browse all the available software project
data. The PROMISE repository of SE data has grown to 200+ projects~\cite{promise15}
and this is just one of over a dozen open-source repositories
that are readily available to researchers~\cite{rod12}.
For example,  the time of this writing (Feb  2015), our web searches show that Mozilla Firefox has over 1.1 million bug reports, and platforms such as GitHub hosts over 14 million projects.

@@

Our tuning results  challenge much prior work in software analytics. 
Firstly, there exist research papers
that use data miners to   show that certain factors
are more influential than others for (say)
predicting defects. As shown below, such conclusions can be dramatically
changed by the tuning process since those  ``influential'' factors are very different pre- and post- tuning. Also, those factors tend to  change from project to project or if the goal
of the tuning is altered.
Hence, many old papers    need to be revisted  and perhaps revised~\cite{bell2013limited,rahman2013how,me02k,moser2008comparative,zimmermann2007predicting,herzig2013predicting}.  
For example, one of us (Menzies) used data miners
to assert that some factors were more important than others for predicting
successful software reuse~\cite{me02k}. That assertion should now be doubted since that
Menzies study did conduct a tuning study before reporting what factors the data miners
found where most influential.

Secondly, several   prominent IEEE TSE papers~\cite{lessmann2008benchmarking,hall11,me07b} have claimed 
that learnerX is better than learnerY for some software analytics task~.
For example, a recent IEEE TSE article claimed that the 
CART decision tree learner was far worse than Random Forests for
software defect prediction~\cite{lessmann2008benchmarking}. 
Such conclusions do not survive tuning.
For example,
after tuning, the worst untuned learner (CART) can out-perform the supposedly
best learner (Random Forests). Hence, all those prior results that ranked learners for software
analytics now need to be revisited and perhaps revised.

Thirdly, it is standard practice to use the default ``off-the-shelf'' tunings  for data mining tools (previously
we have defended that approach on methodological grounds arguing that it
encourages reproducibility~\cite{me15:book1}). That ``off-the-shelf''  policy
can no longer be condoned. For example, one such default setting
  in the Python \mbox{SciKitLearn} toolkit~\cite{scikit-learn}
is to use $F=10$ decision trees in  Random Forest classifiers.
Our optimizer settled on  $F$ values that were were nowhere near that default:

@@

This section offers a small demonstration
of the impact of tuning parameters. It also demonstrates that this issue of tuning effects not jist the complex data miners discussed later in the paper bu t also 
applies for even very simple  learning schemes.

@@

That researcher might test that belief by using studying historical
defect logs with 
linear regression. Here are two equations (learned from the NASA data at goo.gl/pGDfvp)
that use just lines of code or the Halstead measures $N,V,L,D,I,E,B,T$ seen in a
software module (in this case, a ``C'' function).
Note that the Halstead correlation $c_2$
is worse than those from lines of code $c_1$. This result  seems to suggest that
despite their potential support for refactoring, our researchers should not use Halstead.

@@

The rest of this paper repeats the analyis of the last section, but for much
more complex learners. In those results, shown  below, we will find
examples of {\em rank reversals};
i.e.  after tuning the top ranked defector becomes last and vice versa.
Such examples motivate the extensive exploration of tuning in software analytics.
Before presenting those results, this section offers some background notes on defect prediction,
performance measures, and optimization with differential evolution.

@@

The psuedocde for differential evolution is shown in Algorithm \ref{alg:DE}

for 
Differential Evolution (DE)\cite{storn1997differential} is a stochastic search algorithm that 
optimizes a problem by iteratively trying to improve a  population of candidate solutions with 
regard to a given quality measurement. Such method makes no assumptions about the 
problem being optimized and has already been used as a parameter tuner
\cite{omran2005differential, chiha2012tuning}.

DE starts with creating a population of candidate solutions and then generates new 
candidates based on $New = X+f*(Y-Z)$, where $X$, $Y$ and $Z$ are randomly selected 
solutions from the current frontier and $f$ is a crossover factor. The newly generated solution 
will be added into the next generation of solutions if dominating previous old points in the 
frontier. To avoid meaning less iteration, early termination strategy is applied that is at the 
beginning, assign a value to the $life$ parameter, it would be reduced by one each time   
when the new generation of candidate solutions does not improve in terms of quality 
measurement. The DE will stop when the $life$ equals to 0.

Algorithm \ref{alg:DE} is a list of pseudocode of DE with early termination for maximizing a 
score function, where $np$ is the number of population in each generation, $f$ is the 
crossover factor as mentioned, $cf$ is the probability for crossover operation to generate new 
candidate, $life$ is to control termination.

\begin{algorithm}
\caption{Pesudocode for DE with Early Termination}
\label{alg:DE}
\begin{algorithmic}[1]
\Require $np = 10$, $f=0.75$, $cr=0.3$, $life = 5$
\Ensure $S_{best}$

     \Function{DE}{$np$, $f$, $cr$, $life$}
 \State $Population  \gets $ InitializePopulation($np$)
 \State $S_{best} \gets $GetBestSolution($Population $)
 \While{$life > 0$}
\State $NewGeneration \gets \emptyset$
\For{$i=0 \to np-1$}
\State $S_i \gets$ GenNew($Population [i], Population , cr, f$)
\If {Score($S_i$) >Score($Population [i]$)}
\State $NewGeneration$.append($S_i$)
\Else
\State $NewGeneration$.append($Population [i]$)
\EndIf
\EndFor
\State $Population  \gets NewGeneration$
\If{$\neg$ Improve($Population $)}
\State $life -=1$
\EndIf
\State $S_{best} \gets$ GetBestSolution($Population $)
 \EndWhile
\State \Return $S_{best}$
\EndFunction
\Function{Score}{$Candidate$}
   \State set tuned parameters according to $Candidate$
   \State TrainLearner()
   \State $scores \gets$TestLearner()
   \State \% scocres = [pd, f, precision]
   \State $s \gets scores[i] $ 
   \State \% i determines the goal of tuning
   \State \Return$s$
\EndFunction
\Function{GenNew}{$old, pop, cr, f$}
  \State $a, b, c\gets threeOthers(pop,old)$ 
  \State \% pick 3 other points from the population
  \State $newf \gets \emptyset$
  \For{$i=0 \to np-1$}
       \If{$cr < random()$}
         \State $newf$.append($old[i]$)
                \Else
                  \If{typeof(old[i]) == bool}
                    \State $newf$.append(not $old[i]$)
         \Else
          \State $newf$.append(trim($i$,($a[i] + f * (b[i] - c[i]$))))
          \State \%trim() will make new value fall in the range
         \EndIf
       \EndIf
  \EndFor
\EndFunction
        \end{algorithmic}            
\end{algorithm}

@@

\section{Experiment}

\subsection{Data Set}

The data used in this study is from PROMISE repository. Ten software defect predicition data 
sets are analyzed. They're {\it ant}, {\it camel}, {\it ivy}, {\it jedit}, {\it log4j}, {\it lucene}, {\it 
synapse}, {\it velocity}, {\it xalan} and {\it xerces}. Each of these data sets is composed of 
several software modules with number of defects and code attributes. For more detailed 
description of code attributes and the original data sets, please refer to  http://openscience.us/
repo/

\subsection{Experiment Design}

The experiment aims at investigating whether tuning helps learners improve performance in 
terms of accuracy measurement. We choose compare WHERE-based learner with and 
without tuning parameters and two {\it significantly different} learners Random Forest and 
CART according to \cite{lessmann2008benchmarking}. As mentioned above, we'd like to see 
whether tuning will help change the rank of CART and make it comparable with Random 
Forest.

To evaluate accuracy performance of learners, several measurements are proposed to use, like 
probability of  detection({\it pd}) and probability of false alarm ({\it pf})\cite{menzies2007data}, 
the area under the receiver operating characteristics curve ({\it AUC})
\cite{lessmann2008benchmarking}, precision and F measure\cite{zhang2007comments}. In this work, we 
expect learners should identify as many defective modules as possible while avoiding false 
alarm. Therefore, learners are evaluated by both of {\it pd} and {\it pf} simaltanieously. A single 
measure, G-measure, defined as the harmonic mean of {\it pd} and $1-{\it pf}$  is used. The 
G-measure value is between 0 and 1. The higher, the better.
\begin{equation}
G = \frac{2*(1-pf)*pd}{1-pf+pd}
\end{equation}

@@

\begin{figure}[!t]

\renewcommand{\baselinestretch}{0.8}
\scriptsize
\centering
  \begin{tabular}{r@{~}|r@{~}rr|r@{~}rr|r@{~}rr}
    & \multicolumn{3}{c|}{WHERE}&\multicolumn{3}{c|}{CART}& \multicolumn{3}{c}{Random Forest}\\\cline{2-10} 
    Goal & tuned & naive &$\frac{\mathit{tuned}}{\mathit{naive}}$& tuned & naive &$\frac{\mathit{tuned}}{\mathit{naive}}$& tuned & naive&$\frac{\mathit{tuned}}{\mathit{naive}}$\\\hline
pd&71.4&1.2&60&4.4&0.1&44&7.1&0.2&36\\ 
prec&82.8&1.4&59&3.9&0.1&39&7.4&0.2&37\\
f&86.1&1.2&72&3.3&0.1&33&6.7&0.2&34
  \end{tabular}
  \caption{For one train/test set pair (ANT),  time (in seconds) spent tuning,
  for different goals, or just running the ``off-the-shelf'' learner.}
\end{figure}

\begin{figure}[!t]

\renewcommand{\baselinestretch}{0.8}
\scriptsize
\centering
\begin{tabular}{r|rrr}
goal &	WHERE	&CART&	Random Forest\\\hline
f	&15	&13	&9\\ 
pd	&15	&16	&8\\ 
prec&	14	&14	&8
\end{tabular}
\caption{How much slower is tuning vs just running the default
settings of the learner? E.g. bottom right, tuning to increase
precision with Random Forests is eight times slower than
just running Random Forests.}
\end{figure}

@@

\renewcommand{\baselinestretch}{0.8}
\scriptsize
\centering
  \begin{tabular}{c c c c c c c c c c }
  \hline\hline
  Dataset &ant&antV1&antV2&camel&camelV1&ivy&jedit&jeditV1&jeditV2
\\\hline
  training &20/125 &40/178 &32/293 &13/339 &216/608 &63/111 &90/272 &75/306 &79/312
\\  tunning  &40/178 &32/293 &92/351 &216/608 &145/872 &16/241 &75/306 &79/312 &48/367
\\  testing &32/293 &92/351 &166/745 &145/872 &188/965 &40/352 &79/312 &48/367 &11/492
\\  \end{tabular}
   \caption{The percentage of defective instances in each experimental data set. For each experiment,  training, tuning and testing data are composed of single chronological data file}
\end{figure*}
\begin{figure*}[!ht]
\scriptsize
\centering
  \begin{tabular}{c c c c c c c c c c }
  \hline\hline
  Dataset &log4j&lucene&poi&poiV1&synapse&velocity&xerces&xercesV1
\\\hline
  training &34/135 &91/195 &141/237 &37/314 &16/157 &147/196 &77/162 &71/440
\\  tunning  &37/109 &144/247 &37/314 &248/385 &60/222 &142/214 &71/440 &69/453
\\  testing &189/205 &203/340 &248/385 &281/442 &86/256 &78/229 &69/453 &437/588
\\  \end{tabular}

   \caption{The percentage of defective instances in each experimental data set. For each experiment,  training, tuning and testing data are composed of single chronological data file}
\end{figure*}

>>>>

\documentclass{sig-alternative}
\usepackage{multirow}
\usepackage{color}
\usepackage{graphics}
\usepackage{cite}
\usepackage{rotating}
\usepackage{eqparbox}
\usepackage{graphics}
\usepackage{colortbl} 
\usepackage{picture}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\renewcommand{\footnotesize}{\scriptsize}
\definecolor{lightgray}{gray}{0.8}
\definecolor{darkgray}{gray}{0.6}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
%%% graph
\newcommand{\crule}[3][darkgray]{\textcolor{#1}{\rule{#2}{#3}}}
\newcommand{\rone}{\crule{1mm}{1.95mm}}
\newcommand{\rtwo}{\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}}
\newcommand{\rthree}{\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}}
\newcommand{\rfour}{\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}} 
\newcommand{\rfive}{\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}\hspace{0.3pt}\crule{1mm}{1.95mm}}
\newcommand{\quart}[3]{\begin{picture}(100,6)%1
{\color{black}\put(#3,3){\circle*{4}}\put(#1,3){\line(1,0){#2}}}\end{picture}}
\definecolor{Gray}{gray}{0.95}
\definecolor{LightGray}{gray}{0.975}

@@

For the purposes of learning
software defect predictors this optimization strategy can quickly
find  good tunings that  dramatically change   the performance of a learner.
For example,
in this paper we show   tunings that  alter detection  precision  
 from 2\% to 98\%.

These results prompt for a change to standard methods in software analytics.
At least for defect prediction, 
it is no longer enough to just run a data miner and present the result
{\em without} first conducting a tuning optimization study.
The implications for other kinds of software analytics is now an open and pressing questions.

@@

In the $21^{st}$ century, it is now impossible
to manually browse all the available software project
data. The PROMISE repository of SE data has grown to 200+ projects~\cite{promise15}
and this is just one of over a dozen open-source repositories
that are readily available to researchers~\cite{rod12}.
For example,  the time of this writing (Feb  2015), our web searches show that Mozilla Firefox has over 1.1 million bug reports, and platforms such as GitHub host over 14 million projects.

@@

Our tuning results  challenge much prior work in software analytics. 
Firstly, there exist research papers
that use data miners to   show that certain factors
are more influential than others for (say)
predicting defects. As shown below, such conclusions can be dramatically
changed by the tuning process since those  ``influential'' factors are very different pre- and post- tuning. Also, those factors tend to  change from project to project or if the goal
of the tuning is altered.
Hence, many old papers    need to be revisted  and perhaps revised~\cite{bell2013limited,rahman2013how,me02k,moser2008comparative,zimmermann2007predicting,herzig2013predicting}.  
For example, one of us (Menzies) used data miners
to assert that some factors were more important than others for predicting
successful software reuse~\cite{me02k}. That assertion should now be doubted since that
Menzies study did not conduct a tuning study before reporting what factors the data miners
found where most influential.

Secondly, several  prominent IEEE TSE papers~\cite{lessmann2008benchmarking,hall11,me07b} have claimed 
that learnerX is better than learnerY for some software analytics task.
For example, a recent IEEE TSE article claimed that the 
CART decision tree learner was far worse than Random Forests for
software defect prediction~\cite{lessmann2008benchmarking}. 
Such conclusions do not survive tuning.
For example,
after tuning, the worst untuned learner (CART) can out-perform the supposedly
best learner (Random Forests). Hence, all those prior results that ranked learners for software
analytics now need to be revisited and perhaps revised.

Thirdly, it is standard practice to use the default ``off-the-shelf'' tunings  for data mining tools (previously
we have defended that approach on methodological grounds arguing that it
encourages reproducibility~\cite{me15:book1}). That ``off-the-shelf''  policy
can no longer be condoned. For example, one such default setting
  in the Python \mbox{SciKitLearn} toolkit~\cite{scikit-learn}
is to use $F=10$ decision trees in  Random Forest classifiers.
Our optimizer settled on  $F$ values that were nowhere near that default:

@@

This section offers a small demonstration
of the impact of tuning parameters. It also demonstrates that this issue of tuning effects 
not just the complex data miners discussed later in the paper but also 
applies for even very simple  learning schemes.

@@

That researcher might test that belief by using studying historical
defect logs with 
linear regression. Here are two equations (learned from the NASA data at goo.gl/pGDfvp)
that use just lines of code or the Halstead measures $N,V,L,D,I,E,B,T$ seen in a
software module (in this case, a  function).
Note that the Halstead correlation $c_2$
is worse than those from lines of code $c_1$. This result  seems to suggest that
despite their potential support for refactoring, our researchers should not use Halstead.

@@

The rest of this paper repeats the analysis of the last section, but for much
more complex learners. In those results, shown  below, we will find
examples of {\em rank reversals};
i.e.  after tuning the top ranked defector becomes last and vice versa.
Such examples motivate the extensive exploration of tuning in software analytics.
Before presenting those results, this section offers some background notes on defect prediction,
performance measures, and optimization with differential evolution.

@@

\subsubsection{Differential Evolution: The Details}

\input{algo}
The psuedocode for differential evolution is shown in Algorithm~\ref{alg:DE}.
Note that, as we describe the algorithm,
  any superscript number denotes a line in that algorithm.

DE is an evolutionary algorithm; i.e. the next {\em NewGeneration} is learnt from
a current {\em Population}.  If the new is no better than the current, then
we lose one life, terminating when we run out of lives$^5$.

Each candidate solution in the {\em Population}  
is a pair of {\em (Tunings, Scores)}. In this paper, {\em Tunings} are selected from
\fig{parameters} and {\em Scores} come from training a learner using those parameters
and applying it to some test data$^{23-28}$.

The premise of this algorithm is that the best way to mutate existing tunings
is to {\em Extrapolate}$^{29}$
between current solutions.  Three solutions $a,b,c$ are selected at random.
For each tuning parameter $i$, at some probability {\em cr}, we replace  new tuning is created
via:
\bi
\item For numerics: $a_i+f \times (b_i - c_i)$   where $f$ is a parameter
controlling the cross-over amount.  The {\em trim} function$^{39}$ limits the new
value to the legal range min..max of that parameter.
\item For booleans:  negate the current value$^{37}$.
\ei
The main loop of DE$^7$ runs over the {\em Population}, replacing old items
with new {\em Candidate}s (if the new candidate is better than the old item).
This means that, as the loop progresses, the {\em Population} is full of increasiningly
more valuable solutions. This, in turn, also improves  the candidates (which are generated
from the {\em Population}.

For this experiments of this paper, we collect performance
values from a data mining, from which a {\em Goal} function extracts one 
performance value$^{27}$ (so we tun this code many times, each time with
a different {\em Goal}$^2$).  Technically, this makes this a  {\em single objective} DE (and for notes on multi-objective DEs, see~\cite{Coello05,zhang07,5583335}).

@@

\renewcommand{\baselinestretch}{0.8}
\scriptsize
\centering
  \begin{tabular}{c c c c c c c c c c }\hline
  Dataset &antV0&antV1&antV2&camelV0&camelV1&ivy&jeditV0&jeditV1&jeditV2
\\\hline
  training &20/125 &40/178 &32/293 &13/339 &216/608 &63/111 &90/272 &75/306 &79/312
\\  tuning  &40/178 &32/293 &92/351 &216/608 &145/872 &16/241 &75/306 &79/312 &48/367
\\  testing &32/293 &92/351 &166/745 &145/872 &188/965 &40/352 &79/312 &48/367 &11/492
\\  \end{tabular}
   \caption{Ratios of defective instances in each experimental data set. 
   E.g., the top left data set has 20 defective classes out of 125 total.
   This paper runs one experiment for each column
   shown in this figure. The {\em training} set is used by an untuned learner
   to build a model, which is then tested on the {\em testing} data.
   {\em Training} data is used by  differential evolution when it builds
    a model (using one set of possible tunings) and that model is tested on {\em tunings}.
    Finally, the best model from by DE is applied to {\em testing}. 
   This information is continued  in \fig{data2}. 
   }\label{fig:data1}
\end{figure*}
\begin{figure*}[!ht]
\scriptsize
\centering
  \begin{tabular}{c c c c c c c c c c }
  \hline\hline
  Dataset &log4j&lucene&poiV0&poiV1&synapse&velocity&xercesV0&xercesV1
\\\hline
  training &34/135 &91/195 &141/237 &37/314 &16/157 &147/196 &77/162 &71/440
\\  tuning  &37/109 &144/247 &37/314 &248/385 &60/222 &142/214 &71/440 &69/453
\\  testing &189/205 &203/340 &248/385 &281/442 &86/256 &78/229 &69/453 &437/588
\\  \end{tabular}

   \caption{More ratios of  defective instances in each experimental data set. 
   Same format as \fig{data1}.}\label{fig:data2}
\end{figure*}

\section{Experiments}

The following experiments compare the performance of three learners, tuned and untuned, on 17
sets of data.

\subsection{Data Sets}

Our defect data comes from the PROMISE repository\footnote{http://openscience.us/repo}.
It pertains to 
open source Java systems defined in terms of \fig{ck}:  {\it ant}, {\it camel}, {\it ivy}, {\it jedit}, {\it log4j}, {\it lucene}, {\it 
synapse}, {\it velocity}, {\it xalan} and {\it xerces}.

An important principle in data mining is not to test on the data used
in training.  There are many ways to design a experiment that satisfies this principle:
\bi
\item  Leave-one-out experiments can be impractically slow for large data sets;
\item Cross-validation can mix up older and newer data sets so it may well be that
data from the {\em future} is used to train on {\em past data}.
\ei
To avoid these problems, we used an incremental learning approach. The following
experiment ensures that the training data was created at some time before the test
data.

For this experiment, we looked for data sets that had at least three  
releases in PROMISE:
\bi 
\item The {\em first} release was used for some  {\em training}, to collect a baseline
   using an untuned learner.
   \item
   The {\em first} release was also used  on line 24 of Algorithm~\ref{alg:DE} to
   build some model using some the tunings found in some {\em Candidate}.
   \item The {\em second} release was used on 25 of Algorithm~\ref{alg:DE} to 
   test the model found on line 24.
   \item Finally the {\em third} release was used to gather the performance statistics
   reported below from (a)~the model generated by the untuned learner or (b)~the
   best model found by differential evolution.

   \ei
 Note that the above ensures that all treatments get tested on the same final data set.

 Some data sets have more than three releases and, for those data, we could run more
 than one experiment. For example, {\em ant} has five versions in PROMISE so
 we ran three experiments called V0,V1,V2:
 \bi
 \item AntV0: first,second,third = versions 1,2,3
 \item AntV1: first,second,third = versions 2,3,4
 \item AntV2: first,second,third = versions 3,4,5
 \ei 
These data sets are displayed in \fig{data1} and \fig{data2}.

\subsection{Performance Scores}

To evaluate accuracy performance of learners, several measurements are used.
The 
probability of  detection({\it pd}) and probability of false alarm ({\it pf})\cite{menzies2007data}, 
the area under the receiver operating characteristics curve ({\it AUC})
\cite{lessmann2008benchmarking}, precision and F measure\cite{zhang2007comments}. In this work, we 
expect learners should identify as many defective modules as possible while avoiding false 
alarm. Therefore, learners are evaluated by both of {\it pd} and {\it pf} simaltanieously. A single 
measure, G-measure, defined as the harmonic mean of {\it pd} and $1-{\it pf}$  is used. The 
G-measure value is between 0 and 1. The higher, the better.
\begin{equation}
G = \frac{2*(1-pf)*pd}{1-pf+pd}
\end{equation}

@@

@@
